---
title: "hw2"
author: "Yiran Jin"
date: "2026-10-05"
output: github_document
---

load necessary libraries
```{r,message=FALSE,warning=FALSE}
library(tidyverse) 
library(janitor)
```

# Problem 1
```{r}
subway_data = read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") |>     clean_names() |> 
  select(
    line, 
    station_name, 
    station_latitude, 
    station_longitude, 
    starts_with("route"), 
    entrance_type, 
    entry, 
    vending, 
    ada
  ) |> 
  mutate(
    entry = case_match(
      entry, 
      "YES" ~ TRUE, 
      "NO" ~ FALSE 
    ) 
  )
```

## summary of the data
This NYC transit dataset contains information about NYC subway stations and their entrances and exits. The variables include the subway line, station name, station latitude and longitude, routes served, entrance type, whether entry is allowed, whether vending is available, and ADA compliance. I selected the variables which are needed and converted the entry variable from character values (YES and NO) to a logical variable (TRUE and FALSE). The resulting dataset has 1868 rows and 19 columns. These data are not tidy because the routes served by each station are stored in multiple variables rather than having one route per observation. For example, 36th St serves both the N and R trains, which are stored in separate variables (route1 and route2) rather than as separate observations.

## answers of the questions
### distinct stations (only keep the first row of each distinct line and station name)
```{r}
subway_data |> 
  distinct(line, station_name) |> 
  nrow()
```
There are 465 distinct stations.

### ADA Compliant Stations
```{r}
subway_data |> 
  filter(ada == TRUE) |> 
  distinct(line, station_name) |> 
  nrow()
```
There are 84 ADA-compliant stations.

### Entrances/exits Without Vending That Allow Entrance
```{r}
subway_data |> 
  filter(vending == "NO") |> 
  summarize( 
    proportion = mean(entry, na.rm = TRUE) 
  )
```
Among station entrances and exits without vending, approximately 37.7% allow entrance.

## Reformat Route Variables
```{r}
subway_routes =
  subway_data |>
  mutate(
    across(starts_with("route"), as.character)
  ) |>
  pivot_longer(
    cols = starts_with("route"),
    names_to = "route_number",
    values_to = "route_name",
    values_drop_na = TRUE
  ) |>
  filter(route_name != "")
```

## Stations Serving the A Train
```{r}
subway_routes |> 
  filter(route_name == "A") |> 
  distinct(line, station_name) |> 
  nrow()
```
There are 60 distinct stations that serve the A train.

## ADA-Compliant A Train Stations
```{r}
subway_routes |> 
  filter(route_name == "A", ada == TRUE) |> 
  distinct(line, station_name) |> 
  nrow()
```
17 are ADA compliant

# Problem 2
## clean the data--Mr. Trash Wheel
```{r}
library(readxl)

mr_trash_wheel =
  read_excel(
    "data/202509 Trash Wheel Collection Data.xlsx",
    sheet = "Mr. Trash Wheel",
    range = "A2:N709"
  ) |>
  janitor::clean_names() |>
  filter(!is.na(dumpster)) |>
  mutate(
    year = as.integer(year),
    sports_balls = as.integer(round(sports_balls)),
    trash_wheel = "Mr. Trash Wheel"
  )
```

## clean the data--Professor Trash Wheel
```{r}
professor_trash_wheel =
  read_excel(
    "data/202509 Trash Wheel Collection Data.xlsx",
    sheet = "Professor Trash Wheel",
    range = "A2:L134"
  ) |>
  janitor::clean_names() |>
  filter(!is.na(dumpster)) |>
  mutate(
    year = as.integer(year),
    trash_wheel = "Professor Trash Wheel"
  )
```

## clean the data--Gwynnda
```{r}
gwynnda_trash_wheel =
  read_excel(
    "data/202509 Trash Wheel Collection Data.xlsx",
    sheet = "Gwynns Falls Trash Wheel",
    range = "A2:L351"
  ) |>
  janitor::clean_names() |>
  filter(!is.na(dumpster)) |>
  mutate(
    trash_wheel = "Gwynnda"
  )
```

## Combine the datasets
```{r}
trash_wheel =
  bind_rows(
    mr_trash_wheel,
    professor_trash_wheel,
    gwynnda_trash_wheel
  )
dim(trash_wheel)
```

## Summary of the data
The Trash Wheel dataset contains information on dumpsters collected by three trash wheels: Mr. Trash Wheel, Professor Trash Wheel, and Gwynnda. Key variables include the dumpster number, month, year, date, weight and volume of trash collected, and counts of different types of trash such as plastic bottles, polystyrene, cigarette butts, plastic bags, and wrappers. I cleaned each dataset by selecting the appropriate worksheet and data range, removing rows that did not contain dumpster-specific observations, and using reasonable variable names. For Mr. Trash Wheel, I also rounded the number of sports balls to the nearest integer and converted it to an integer variable. I changed 'years' column to integer for each dataset in order to combine three datasets. Finally, I added a variable identifying the trash wheel and combined the three datasets into one tidy dataset. The resulting dataset contains `r nrow(trash_wheel)` observations.

## total weight of trash collected by Professor Trash Wheel
```{r}
professor_trash_wheel |>
  summarize(
    total_weight = sum(weight_tons, na.rm = TRUE)
  )
```
Professor Trash Wheel collected a total of 282.26 tons of trash.

## total number of cigarette butts collected by Gwynnda in June of 2022
```{r}
gwynnda_trash_wheel |>
  filter(
    month == "June",
    year == 2022
  ) |>
  summarize(
    total_cigarette_butts = sum(cigarette_butts, na.rm = TRUE)
  )
```
Gwynnda collected a total of 18,120 cigarette butts in June 2022

# Problem 3
```{r}
baseline =
  read_csv(
    "data/MCI_baseline.csv",
    skip = 1,
    na = "."
  ) |>
  rename(
    study_id = ID,
    age = `Current Age`,
    sex = Sex,
    education = Education,
    apoe4 = apoe4,
    mci_onset_age = `Age at onset`
  ) |>
  mutate(
    study_id = as.integer(study_id),
    age = as.numeric(age),
    sex = case_match(
      sex,
      1 ~ "Male",
      0 ~ "Female"
    ),
    apoe4 = case_match(
      apoe4,
      1 ~ "Carrier",
      0 ~ "Non-carrier"
    ),
    education = as.numeric(education),
    mci_onset_age = as.numeric(mci_onset_age)
  ) |>
  filter(
    is.na(mci_onset_age) | mci_onset_age > age
  )
```

## How many participants were recruited
```{r}
baseline |>
  nrow()
```
479 participants are recruited.
