---
title: "hw2"
author: "Yiran Jin"
date: "2026-10-05"
output: github_document
---

load necessary libraries
```{r}
library(tidyverse) 
library(janitor)
```

# Problem 1
```{r}
subway_data = read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") |>     clean_names() |> 
  select(
    line, 
    station_name, 
    station_latitude, 
    station_longitude, 
    starts_with("route"), 
    entrance_type, 
    entry, 
    vending, 
    ada
  ) |> 
  mutate(
    entry = case_match(
      entry, 
      "YES" ~ TRUE, 
      "NO" ~ FALSE 
    ) 
  )
```

## summary of the data
This NYC transit dataset contains information about NYC subway stations and their entrances and exits. The variables include the subway line, station name, station latitude and longitude, routes served, entrance type, whether entry is allowed, whether vending is available, and ADA compliance. I selected the variables and converted the entry variable from character values (YES and NO) to a logical variable (TRUE and FALSE). The resulting dataset has 1868 rows and 19 columns. These data are not fully tidy because the routes served by each station are stored in multiple variables (route1 through route11) rather than having one route per observation.

